---
title: "Graphical Data Analysis"
author: "Paul Collett"
date: '2020-11-27'
output: 
  rmdformats::downcute:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    df_print: paged
    code_folding: show
  html_notebook:
    df_print: paged
    toc: true
    toc_float: true
    number_sections: true
  word_document:
    toc: true
  bibliography: bibliography.bib
---
```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, class.source = "bg-success", fig.height=10, fig.width=10)
```
```{r}

```


```{css, echo = FALSE}
p, li {
  font-size: 24px;
  line-height: 1.5;
  margin-top: 24px;
  margin-bottom: 12px;
}
img {
  margin-top: 15px;
  margin-bottom: 30px;
}
.bg-success {
  font-size: 18px;
  margin-bottom: 18px;
}
h4.author {font-size: 24px;}
.caption {
  font-size: 18px;
  font-style: italic;
}
#note {
  padding: 5px;
  background: #FF9300;
  color: white;
  }

```
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include = FALSE}
library(tidytext)
library(tidyr)
library(dplyr)
library(ggplot2)
library(hrbrthemes)
library(ggExtra)
library(reshape2)
library(ggridges)
library(dabestr)
library(readtext)
library(quanteda)
library(truncnorm)
library(cowplot)
library(ggstatsplot)
library(viridisLite)
library(rmdformats)
```
# Preparing your data

Set your working directory where you will store the R script and any other files you may need for the project:

```{r}
setwd("~/Documents/R Worksets/JALT2020 Workshop")
```

Create bimodal distribution data set to illustrate normality assumptions:

```{r}
nn <- 100
set.seed(1234)
sim1 <- c(rtruncnorm(nn/2, a=0, b=10, mean=2, sd=.75),
          rtruncnorm(nn/2, a=0, b=10, mean=8, sd=.75))
```

Create a more normally-distributed second sample:

```{r}
set.seed(1234)
sim2 <- rnorm(100, mean = 5, sd = .75)

x1 <- as.data.frame(sim1)
x2 <- as.data.frame(sim2)
```

Store the data in a format (dataframe) that can easily be retrieved for the analysis:

```{r}
x3 <- as.data.frame(cbind(sim1, sim2)) ## this joins the two data sets together and transforms them into a dataframe
write.csv(x3, "x3_data.csv") ## this command writes the dataframe to a csv file ("x3_data.csv") and saves it on your computer in the working directory
```

You can check there are no problems with the data by using the `head` & `tail` commands which display the first & last few lines respectively of the data:
  
```{r}
head(x3); tail(x3)
```

## Histograms to show data distribution.

Using the `hist` command which is part of the base installation of R. Work with the `sim1` & `sim2` samples from above. 

```{r fig.height=5, fig.width=8}
hist(sim1, prob=F, main = "", xlab = "Group 1", xlim = c(0,10), ylim = c(0,20), breaks=15, cex.main=1.5, cex.lab=1.5, cex.axis=1.5)
abline(v=mean(sim1),col="blue", lty = 2)

hist(sim2, prob=F, main = "", xlab = "Group 2", xlim = c(0,10), breaks=15, cex.main=1.5, cex.lab=1.5, cex.axis=1.5)
abline(v=mean(sim1),col="blue", lty = 2)
```

## QQ Plots.

Using the `qqnorm` command which is part of the base R installation. Work with the `sim1` & `sim2` samples from above.

```{r}
qqnorm(sim1, cex.main=1.5, cex.lab=1.5, cex.axis=1.5, main = "Group 1")	# this produces the graph
qqline(sim1) # this adds a line showing a theoretical normal distribution for comparison

qqnorm(sim2, cex.main=1.5, cex.lab=1.5, cex.axis=1.5,  main = "Group 2")	# this produces the graph
qqline(sim2)  # this adds a line showing a theoretical normal distribution for comparison
```

# Boxplots and variants for comparisions

**Boxplots** are good for showing the difference between two (or more) samples, e.g. when you would do a t-test or an ANOVA (or a non-parametric equivalent). 

* To create a basic boxplot using base R installation commands:
  
```{r  fig.height=5, fig.width=12}
boxplot(sim2, col = "white", ann = F, horizontal = T, ylim = c(3,7))
#using dataset sim2, no annotations on the axes, y axis range is from 3 to 7
```


## Boxplots for simulated data set 1

These are created using the ggplot2 package. This can be installed and activated as follows:

```{r eval = FALSE}
install.packages("ggplot2")
library(ggplot2)
```

The code to generate the boxplot is more complex here. The first three lines of code create the plot; the following code handles the appearance of the plot. In this case we're creating two plots, fig5a & fig5b.


```{r}
fig5a <- ggplot(x1) +
  aes(x = "", y = sim1) +
  geom_boxplot() +
  # These commands control the appearance of aspects of the theme
  theme_minimal() +
  theme(axis.text.x = element_text(size=14), 
        axis.text.y = element_text(size = 14),
        axis.title = element_text(size = 16, lineheight = 2),
        plot.title = element_text(size = 14, lineheight = 2, face="bold"), 
        strip.text.x = element_text(size = 14)) +
  ylim(0, 10) +
  labs(x = "Group 1", y = "Comprehension Score")

fig5b <- ggplot(x2) +
  aes(x = "", y = sim2) +
  geom_boxplot() +
  theme_minimal()  +
  theme(axis.text.x = element_text(size=14),
        axis.text.y = element_blank(),
        axis.title = element_text(size = 16, lineheight = 2),
        plot.title = element_text(size = 14, lineheight = 2, face="bold"), 
        strip.text.x = element_text(size = 14)) +
  ylim(0, 10) +
  labs(x = "Group 2", y = "")
```

Load the `cowplot` package. This lets you plot different graphs on the same grid...

```{r eval = FALSE}
install.packages("cowplot")
library(cowplot)
```

...using the `plot_grid` command from `cowplot`

```{r}
plot_grid(fig5a, fig5b)
```



## Boxplots for simulated data set 1 augmented with jittered data

Here we add the data points to the boxplot. This involves just one extra line of code:


```{r fig.cap = "Figure 6"}
fig6a <- ggplot(x1) +
  aes(x = "", y = sim1) +
  geom_boxplot() +
  # This additional line of code adds the data points
  geom_jitter(width = .2, size = 3, colour = "orange") + 
  theme_minimal() +
  theme(axis.text.x = element_text(size=14),
        axis.text.y = element_text(size = 14),
        axis.title = element_text(size = 16, lineheight = 2),
        plot.title = element_text(size = 14, lineheight = 2, face="bold"), 
        strip.text.x = element_text(size = 14)) +
  ylim(0, 10) +
  labs(x = "Group 1", y = "Comprehension Score")

fig6b <- ggplot(x2) +
  aes(x = "", y = sim2) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = .2, size = 3, colour = "blue") +
  theme_minimal()  +
  theme(axis.text.x = element_text(size=14), 
        axis.text.y = element_blank(),
        axis.title = element_text(size = 16, lineheight = 2),
        plot.title = element_text(size = 14, lineheight = 2, face="bold"), 
        strip.text.x = element_text(size = 14,)) +
  ylim(0, 10) +
  labs(x = "Group 2", y = "")
plot_grid(fig6a, fig6b)
```

## Dotplot for simulated data set 1

An alternative way to display the data is as a dotplot

```{r}
fig7a <- ggplot(x1) + 
  aes(x = "", y = sim1, fill = "sim1") +
  # main changes are as follows
  stat_summary(fun = median, fun.min = median, fun.max = median, geom = "crossbar", width = 0.6, size = 0.4, color = "black", alpha = 0.6) + # generate the bar showing the median score
  geom_dotplot(binaxis ="y", binwidth = 0.2, stackdir = "center", stackratio = 1.5) + # plot the data as individual points 
  theme_minimal() +
  theme(legend.position = "none", 
        axis.text.x = element_text(size=14),
        axis.text.y = element_text(size = 14),
        axis.title = element_text(size = 16, lineheight = 2),
        plot.title = element_text(size = 14, lineheight = 2, face="bold"), 
        strip.text.x = element_text(size = 14)) +
  ylim(0, 10) +
  labs(x = "Group 1", y = "Comprehension Score")

fig7b <- ggplot(x2) + 
  aes(x = "", y = sim2, fill = "sim1") +
  stat_summary(fun = median, fun.min = median, fun.max = median, geom = "crossbar", width = 0.6, size = 0.4, color = "black", alpha = 0.6) + # generate the bar showing the median score
  geom_dotplot(binaxis ="y", binwidth = 0.2, stackdir = "center", stackratio = 1.5) +
  theme_minimal() + 
  theme(legend.position = "none", 
        axis.text.x = element_text(size=14), 
        axis.text.y = element_blank(), 
        axis.title = element_text(size = 16, lineheight = 2),
        plot.title = element_text(size = 14, lineheight = 2, face="bold"),
        strip.text.x = element_text(size = 14)) +
  ylim(0, 10) +
  labs(x = "Group 2", y = "")

plot_grid(fig7a, fig7b)
```

## Violin plots for simulated data set 1

Violin plots show how the data is distributed and as such are helpful for understanding the structure of your dataset. 

```{r}
fig8a <- ggplot(x1) +
  aes(x = "", y = sim1) +
  # This is the only change from figure 5, calling for a violin plot rather than a boxplot.
  # The adjust and scale arguments set the size of the plots.
  geom_violin(adjust = 1L, scale = "count") + 
  theme_minimal() +
  theme(axis.text.x = element_text(size=14),
        axis.text.y = element_text(size = 14),
        axis.title = element_text(size = 16, lineheight = 2),
        plot.title = element_text(size = 14, lineheight = 2, face="bold"), 
        strip.text.x = element_text(size = 14)) +
  ylim(0, 10) +
  labs(x = "Group 1", y = "Comprehension Score")

fig8b <- ggplot(x2) +
  aes(x = "", y = sim2) +
  geom_violin(adjust = 1L, scale = "area") +
  theme_minimal() +
  theme(axis.text.x = element_text(size=14),
        axis.text.y = element_blank(),
        axis.title = element_text(size = 16, lineheight = 2),
        plot.title = element_text(size = 14, lineheight = 2, face="bold"), 
        strip.text.x = element_text(size = 14)) +
  ylim(0, 10) +
  labs(x = "Group 2", y = "")

plot_grid(fig8a, fig8b)
```

## Notched boxplots

```{r}
set.seed(1234)
response = rnorm(n = 80, mean = c(74, 70), sd = c(3, 4.5))
group = rep(letters[1:2], length.out = 80)
sim4 <- data.frame(group,
                   response)

ggplot(sim4) +
  aes(x = group, y = response) +
  geom_boxplot(notch = TRUE, notchwidth = 0.75) +
  theme_minimal()  +
  theme(axis.text.x = element_text(size=14), # These commands control aspects of the theme, in this case the axis text
        axis.text.y = element_text(size = 14),
        axis.title = element_text(size = 16, lineheight = 2),
        plot.title = element_text(size = 14, lineheight = 2, face="bold"), 
        strip.text.x = element_text(size = 14)) +
  ylim(55, 85) +
  labs(x = "Groups", y = "")
```

# Scatterplots - Looking at Relationships

## Anscombe’s scatterplots.
(Code taken from the _dataset_ package - see https://rdrr.io/r/datasets/anscombe.html)

Visualising the dataset doesn't give a lot of information

```{r}
anscombe
```

Printing a summary of the data suggests each dataset is very similar:

```{r}
summary(anscombe)
```

```{r include = FALSE}
ff <- y ~ x
mods <- setNames(as.list(1:4), paste0("lm", 1:4))
for(i in 1:4) {
  ff[2:3] <- lapply(paste0(c("y","x"), i), as.name)
  ## or   ff[[2]] <- as.name(paste0("y", i))
  ##      ff[[3]] <- as.name(paste0("x", i))
  mods[[i]] <- lmi <- lm(ff, data = anscombe)
  print(anova(lmi))
}

sapply(mods, coef)
lapply(mods, function(fm) coef(summary(fm)))
```

As do the results of a regression analysis comparing the datasets:

```{r}
summary(lm(y1 ~ x1, dat = anscombe))
summary(lm(y2 ~ x2, dat = anscombe))
summary(lm(y3 ~ x3, dat = anscombe))
```


```{r include = FALSE}

op <- par(mfrow = c(2, 2), mar = 0.1+c(4,4,1,1), oma =  c(0, 0, 2, 0))
```

However, graphing the data shows how the statistics can be misleading:

```{r}

for(i in 1:4) {
  ff[2:3] <- lapply(paste0(c("y","x"), i), as.name)
  plot(ff, data = anscombe, col = "red", pch = 21, bg = "orange", cex = 1.2,
       xlim = c(3, 19), ylim = c(3, 13))
  abline(mods[[i]], col = "blue")
}
```





## Scatterplot with loess and regression lines for simulated dataset 3

First, generate the dataset to use

```{r}
sim5 <- data.frame(Subject = (1:30),
                   Group = rep(c("A","B")),
                  x = c(1,1,1,2,2,2,3,3,4,4,4,5,5,5,5,5,6,6,6,6,6,7,7,7,7,8,8,8,9,9),
                  y = c(6,3,5,1,9,4,6,2,11,4,12,7,13,6,10,6,18,4,17,7,16,7,10,6,14,5,15,9,16,12))

group_names <- c(
  A = "Group A",
  B = "Group B"
)
```
Next, generate the plot. This includes a regression line (dotted blue line) with confidence interval band, and a loess line (the curved line). 

```{r}
ggplot(sim5) +
  aes(x = x, y = y) +
  geom_point(size = 3, aes(shape = Group, colour = Group), alpha = .8) +
  geom_smooth(span = 0.75, se = F, aes(colour = Group, linetype = "dashed")) +
  geom_smooth(method = "lm", aes(fill = Group, linetype = "dotted"), alpha = 0.1) +
  theme_minimal() +
  scale_colour_viridis_d(begin = .3, end = .7) +
  scale_fill_viridis_d(option = "C", begin = .2, end = .8) +
  theme(axis.text.x = element_text(size=14),
        axis.text.y = element_text(size = 14),
        axis.title = element_text(size = 16, lineheight = 2),
        plot.title = element_text(size = 14, lineheight = 2, face="bold"),
        strip.text.x = element_text(size = 14)) + 
  scale_linetype_discrete(guide = FALSE) 
```

<<<<<<< HEAD
# Estimation Plots

=======
<<<<<<< HEAD
# Estimation Plots

=======
>>>>>>> 44d11e9491b3013d93274388267885992afc6576
>>>>>>> 56d3c09d8caa9731eaeb42dcaa32b5407f28fbb9
## Unpaired estimation plot (2 sample)

An alternative to carrying out statistical significance tests. 
First, transfer the data set to the correct format

```{r}
sim5.long <- reshape2::melt(sim5, id = c("Subject", "Group"), measured = c("x", "y"))
```

```{r}
fig12 <- dabest(sim5.long, variable, value,
                             idx = c("x", "y"),
                             paired = FALSE) %>% 
  cohens_d()

fig12

plot(fig12, color.column = Group, rawplot.ylabel = "Group Scores", effsize.ylabel = "Effect Size (Cohens d) and 95% CI (5,000 bootstrap resamples)")
```

## Paired estimation plot (2 sample)

```{r}
fig13 <- dabest(sim5.long, variable, value,
                           idx = c("x", "y"),
                           paired = TRUE, id.col = Subject)
fig13.effect <- cohens_d(fig13)  


plot(fig13.effect, color.column = Group, rawplot.ylabel = "Group Scores", effsize.ylabel = "Effect Size (Cohens d) and 95% CI (5,000 bootstrap resamples)")

```


# Extras

The following was not included in the original paper due to word count limitations. It is presented here as a supplement.

## ggstatsplot

This [package](https://indrajeetpatil.github.io/ggstatsplot/index.html) plots graphs and supplements them with details of corresponding statistical tests, providing a streamlined way to test data and visualize the results. 

The example below is generated from the simulated dataset used earlier for the notched boxplots, showing the results of a robust (Yuen) between-group _t_-test with an explanatory measure of effect size, and means and outliers annotated on the graph. Note that the CIs listed here are different to those in the paper. The CIs reported here are for the effect size in the graph, while those in the paper are the CIs of the difference in means. Other options for the type of _t_-test used are possible.

```{r}
ggstatsplot::ggbetweenstats(
  data = sim4,
  x = "group",
  y = "response",
  outlier.tagging = TRUE,
  type = "robust"
  )
```

Here is another example, a variation of the scatterplot made using the sim5 dataset:

```{r}
ggstatsplot::ggscatterstats(
  data = sim5,
  x = x,
  y = y,
  xlab = "X Variable",
  ylab = "Y Variable",
  title = "An example scatterplot with ggstatsplot",
  type = "nonparametric"
)
```

This produces the scatterplot with additional _marginal plots_ showing the distribution of the X & Y variables. The results of a _t_-test, and a non-parametric correlation coefficient are included. 

The `ggstatsplot` package offers a lot of options. By incorporating functions from numerous other packages, it makes generating and displaying results somewhat easier than if using each package separately. This is something that should be helpful for understanding your results when carrying out an analysis.


## Other options

* The data we will work with here is available on my GitHub repository https://github.com/pcjapan/graphical-data-analysis. The first data set is called [_comprehension-data.txt_](https://raw.githubusercontent.com/pcjapan/graphical-data-analysis/2c7d658e59f78610d2cc857b62eb79c3f4461b2f/comprehension-data.txt). Either load the data directly from GitHub or save a copy of the data in your working directory, and then read into R: 

```{r}
working <- read.delim("https://raw.githubusercontent.com/pcjapan/graphical-data-analysis/2c7d658e59f78610d2cc857b62eb79c3f4461b2f/comprehension-data.txt", header = TRUE, sep = "\t")
```

* To work with the `ggplot2` library, and many other functions in R, your data has to be in `long` format, as opposed to `wide`
* `Wide` format is where each individual subject is recorded as a unique row of your excel/csv file, with a column for every measured variable related to that subject, `Long` format has every row containing a measure for a particular variable for each subject. In `wide` data the subjects will not be repeated in rows. In `long` data, the subject may be listed in multiple rows as the data is grouped by the subject response to each variable under study.
* It's fairly easy to change the format of your data in R. One way is by using the `melt` function in the `reshape2` library (this is part of the R base installation):

```{r}
require(reshape2)
```

then run the necessary code. Remember, we're working with data that has been loaded into the "working" dataframe, and we will save the transformed data into a new dataframe named "longComp".


```{r}
longComp <- melt(working, id = c("ID", "Class"), measured = c("Pretest", "Posttest"))
names(longComp) <- c("Student", "Class", "Test", "Comprehension")
```

To see the difference in the data, look at the first few rows.

Here, we have a pretest and posttest score for each student from separate classes on a comprehension test. Scores are entered as one row for each student.

```{r}
head(working) ## Original data
tail(working)
```

Now the data has been rearranged to list in each row the comprehension score for each student by the kind of test. 

```{r}
head(longComp) ## New data
tail(longComp)
```

For any research project, first check your data to see how it is distributed

* Check data for normality - generate a qqplot

```{r fig.height=5, fig.width=8}
qqnorm(working$Pretest, pch = 1, frame = FALSE)
qqline(working$Pretest, col = "steelblue", lwd = 2)
```

* Or look at a histogram of the distribution.

Here, think about what you are expecting to see in the data. A skewed distribution may be what your theory predicts, so may not be a problem; however, it raises issues for the kinds of statistical tests you can carry out.

```{r}
ggplot(working, aes(x=Pretest)) + 
  geom_histogram(aes(y=..density..), colour="gray", fill="white", binwidth = .3)+
  geom_density(color = "red", alpha=.2, fill="#FF6666") 
```


**Line graphs** can show changes in trends in a dataset, but are limited in the information they show. Augmenting with confidence interval bars makes them slightly more informative. Here, we use the `facet_grid` function in `ggplot2` to plot the three graphs on individual panels and display them side-by-side, rather than having them all plotted on the same panel which would reduce readability.  

```{r fig.height=5, fig.width=10}
line <- ggplot(longComp, aes(Test, Comprehension, color = Class))
line + stat_summary(fun = mean, geom = "point") + 
  stat_summary(fun = mean, geom = "line", aes(group = Class)) +
  theme_minimal() + 
  stat_summary(fun.data = mean_cl_boot, geom = "errorbar", width = 0.2) + ## Add confidence interval (CI) error bars
  facet_grid(.~Class)
```

## Theming graphs

Here is where the power of `ggplot` comes into play. You can create a more attractive version of the above or modify most aspects of the display by setting various paramater

* If you haven't already done so, install (you only need to do this the first time you use it) and load the current version of the ggplot library, `ggplot2`:

```{r eval = FALSE}
install.packages(ggplot2)
```

* Then load the package into R. You need to do this with all packages you wish to use:

```{r eval = FALSE}
library(ggplot2) #load and activate the package
```

* Generate the graph, here we're making a boxplot:

```{r}
boxplot <- ggplot(longComp, aes(Class, Comprehension)) #this creates the base graph

CompBox <- boxplot +
  geom_boxplot(outlier.shape = NA, aes(fill = Class)) +  # adding layers to the graph - this sets up the boxplot, using the "Class" condition to identify the data 
  labs(x = "Class", y = "Mean Comprehension Score") + # add labels to the x & y axes
  theme(legend.position = "none", axis.text.x = element_text(size=14), # These commands control the appearance of aspects of the theme
        axis.text.y = element_text(size = 14),
        axis.title = element_text(size = 16, lineheight = 2),
        plot.title = element_text(size = 14, lineheight = 2, face="bold"), 
        strip.text.x = element_text(size = 14), 
        panel.background = element_rect(fill = "snow1")) + 
  viridis::scale_fill_viridis(discrete = TRUE, alpha=0.6) + # Use the viridis colour palettes for colour choices - improves readability
  ggtitle("Changes in Comprehension Scores") + # add a title
  scale_y_continuous(breaks=seq(0,10,1)) + # set the y-axis scale
  facet_wrap(~Test) # this creates a panelled view where the graphs are displayed in separate panels based on the "Test" condition

CompBox  +
  geom_jitter(color="black", size=2, alpha=0.6, width = 0.1) #here you are adding the data points to the graph
```

**Violin plots**  To get a violin plot, just change a couple of lines in the code above:

```{r}
violinplot <- boxplot +
  geom_violin(scale = "count", aes(fill = Class)) + # adding the plot layer to the graph - "count" means areas are scaled proportionally to the number of observations
  geom_boxplot(width=0.1, color="grey60", alpha=0.8, ) + # optionally adding boxplots to provide another layer of detail
  labs(x = "Class", y = "Mean Comprehension Score") + 
  theme(legend.position = "none", axis.text.x = element_text(size=14), 
        axis.text.y = element_text(size = 14),
        axis.title = element_text(size = 16, lineheight = 2),
        plot.title = element_text(size = 14, lineheight = 2, face="bold"), 
        strip.text.x = element_text(size = 14), 
        panel.background = element_rect(fill = "snow1")) + 
  viridis::scale_fill_viridis(discrete = TRUE, alpha=0.6) +
  ggtitle("Changes in Comprehension Scores") + 
  scale_y_continuous(breaks=seq(0,10,1)) + 
  facet_wrap(~Test) 

violinplot
```

## Visualising Correlations

R has strong graphic support for visualising correlations. Let's look at this with a new set of data, corr_Dat. This data was from a study into developing a scale to measure certain attitudes of learners towards factors that influenced approach to language study. The file is once again available in GitHub: [_correlation.txt_](https://github.com/pcjapan/graphical-data-analysis/blob/da29f849d1bc8891e4c1e9413a23d1777e9a6196/correlation.txt)    

* Base R

While this doesn't show you the exact correlations, it provides a scatterplot that shows how each variable is related

```{r}
corr_Dat <- read.csv("https://raw.githubusercontent.com/pcjapan/graphical-data-analysis/da29f849d1bc8891e4c1e9413a23d1777e9a6196/correlation.txt", head = TRUE)
pairs(corr_Dat)
```
* Using the `psych` library

Using the `psych` library `pairs.panels` function is more informative:

```{r}
psych::pairs.panels(corr_Dat, main="Pretest - Class C", method = "spearman")
```
* Or use the `corrplot` library to visualise correlation matrices, as in the following example.

```{r}
cPre <- cor(corr_Dat, method = "spearman") # create correlation matrix
corrplot::corrplot.mixed(cPre)
```

## Ridgeline Plots

A **ridgeline** plot shows the distribution of a numeric variable for a number of groups. Helpful for likert scale results, for example. Use the `ggridges` library along with `ggplot` to create this particular kind of graph.

In this example, we will use one more dataset, which contains the mean scores for subscale respones from the same scale used for the correlation example above. Again, get it from GitHub: the file is [_scale_means.txt_](https://github.com/pcjapan/graphical-data-analysis/blob/b2cd63b9eea32af2d5e3a26c0fd25feacb5011fa/scale-means.txt)

```{r}
library(ggridges) # Remember to load ggplot2 if you haven't already, too
scaleMeans <- read.delim("https://raw.githubusercontent.com/pcjapan/graphical-data-analysis/JALT2020/scale-means.txt", header = TRUE, sep = "\t")
```

R cannot tell which order to treat factors, so puts them into alphabetical order when it runs commands. This means that graphs, etc, may display things in a different order to which you want. In this case, the factor _Test_ has two levels, _Pre_ and _Post_. If we create the graph as is, the _Post_ test results will be displayed before the _Pre_ results, which could be confusing. To fix this, you can easily relevel factors to put them into the correct order for the analysis:

```{r}
scaleMeans$Test <- factor(scaleMeans$Test, levels = c("Pre", "Post")) 
```

Then we can run the code to output the graph:

```{r}
ridgeG <- ggplot(scaleMeans, 
                 aes(x = MResponse, y = Scale, fill = Scale)) # This generates the basic graph

ridgeG + 
  geom_density_ridges2(alpha = 0.8) + 
  facet_wrap(~Class + Test, ncol = 2) + 
  theme_ridges() + 
  scale_x_continuous(expand = c(0.01, 0)) + 
  scale_y_discrete(expand = c(0.01, 0)) + 
  viridis::scale_fill_viridis(discrete = TRUE, alpha=0.6) + 
  theme(legend.position = "none") +
  xlab("Mean Response") # Theming, adding some colour, and separating the graphs onto individual panels
```

# End

That covers everything for now. Thank you for your interest.



